
% Distributed under the MIT License.
% (See accompanying file LICENSE.txt)
% (C) Copyright NoWork team

\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{dirtree}

\lstset{language=ML}

\title{NoWork\\
Developer manual}
\author{NoWork development team\\[2em]}
\date\today

\begin{document}
\maketitle

\section{Introduction}
%% Pierre
Briefly present the project and say which preliminaries document the dev should have read first (wiki, user-manual, coding-style,...).

\section{Installing and hack on NoWork}
%% Matthieu
How to install and begin to develop with NoWork.

\section{Architecture}
%% Mathieu

Here comes a simplified description of the directory layout :
\bigskip

\dirtree{%
.1 /.
.2 doc.
.2 data.
.3 error.
.3 test.
.4 run-pass.
.4 run-fail.
.2 src.
.3 algorithms.
.3 auto\_gen.
.3 generator.
.3 interactive.
.3 parser.
.3 system.
}
\bigskip

And now a more precise description of these directories :

\begin{itemize}

\item \emph{doc} : obviously contains documentation files (user manual, developper manual, coding style, ...)
\item \emph{data/error} : custom exceptions declaration in \emph{*.conf} files (need to be declared here to be used in tests)
\item \emph{data/test} : functional tests framework, test files are added in either \emph{run-fail} or \emph{run-pass}, and test expectations in \emph{test.nw}
\item \emph{src/algorithms} : general algorithms that don't rely on one particular system such as matching and rewriting
\item \emph{src/generator} : generated errors system
\item \emph{src/interactive} : top-level system
\item \emph{src/system} : contains the different AST representations, type checking modules (\emph{type\_checking} and \emph{term\_checker}), and symbol tables

\end{itemize}

\section{Top-level}
%% Vincent
Present the top-level and coding specificities.

\section{Parsing and data representation}

\subsection{System representation}
%% Rémy
Which transformation of the representation of the system occurred and why?

\subsection{Term representation}
%% Pierrick Matthieu
During the process between the parsing and the rewriting of a term, the
structure reprensenting the term is transformed step by step.

\subsubsection{From parsing to semantic-checking}
Parsed terms are transformed in the following \verb?term_ast? :
\begin{lstlisting}
  type info = Lexing.position

  type ident = string

  type term_desc =
  | Const
  | Term of term_ast list
  | Var

  and term_ast =
  {
    name : string;
    info : info;
    desc : term_desc;
  }
\end{lstlisting}

This AST is used only to keep position (in the original file) and structure
informations.
But it is made using only grammatical informations, so it can be
semantically incoherent. That is why it will be transformed during the
type-checking.

\subsubsection{From semantic-checking to type-checking}
The second transformation take a \verb?term_ast? and use semantic informations
contained in the rewriting system to check the semantic of the AST (term
arity and well definition of each symbol)  then
construct an AST with that additional informations (differentiation between
bounded atoms and atom's binders). Now, the term is ready to be type-checked.
\\
N.B. : This AST is called \verb?term_ast_with_binders? (because of his
author's lack of inspiration) in the \verb?term_ast_typed.ml?.
It constructed by \verb?construc_ast_checked? in \verb?term_checker.ml?

\subsubsection{From type-checking to hash-consing}
At the end, the type-checking, processed on the precedent AST, gives a
typed AST keeping only structure and type informations of the original AST.
It is made by \verb?check_type_of_term?.
\\
A term passing all of the steps is considered as wel formed and typed, then
the user can manipulate and particulary rewrite it. In order to have an
efficient rewriting the AST must be transformed a last time.

\subsubsection{Hash-consing}
%% Pierrick
\input{dev_hashconsing.tex}

\section{Type-checking}
%% Ma(t.)hieu

\section{Algorithms}

\subsection{Pattern-matching}
%% Pierre

\subsection{Rewriting}
%% Roven

\section{Rational}
%% Everybody
Important decision that we made such as : Why have we split the language in an interactive and "normal" mode? Why do we have only one parser/lexer (instead of one for system and one for term).


\end{document}
